On the Functionality of the Multilayer Perceptron

### 2020-09-26 Thesis/Drafts/Buell-PHYS799-ThesisDraft0.tex
Added Full Thesis Draft v0

% ================================
% Landon Buell
% Kevin Short
% Physics 799%
% 30 July 2020 
% ================================

@@

@@

@@

\section{The Neural Network}
\label{sec-TheNeuralNetwork}

@@

\subsection{An Introduction to Neural Networks}
\label{sec-NerualNetworkIntro}

@@

\subsection{The Structure of a Neural Network}
\label{sec-NetworkStructre}

\paragraph*{}A Neural Network is simply a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. It is this process of transforming inputs successively in a particular order until an output is attained \cite{Geron,Loy}. This output encodes the models final "decision" given a unique input.

@@

\paragraph*{}The model is recursive by nature, the activations from one layer, $l-1$, are used to directly produce the activations of the next successive layer $l$. Thus eqn. (\ref{eqn-LayerFunction}) can be alternatively written as:
\begin{equation}
\label{eqn-altLayerFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R} \rightarrow x^{(l)} \in \mathbb{R}
\end{equation}
The array of activations, $x^{(0)}$, is the raw input given the neural network, most commonly called \textit{features}. Conversely, the activations $x^{(L-1)}$ are commonly called the network \textit{output} \cite{Geron,James,Loy}.

\begin{algorithm}[H]
\caption{Forward propagation system in a standard deep neural network. Each iteration in the main \textit{for-loop} represents the execution of a layer, and passing the result to the "next" layer function. A practical application of this algorithm should include batches of samples instead of a single sample.
\textcolor{red}{I will put this some place better soon! It is here for future placement}}
\label{algFeedForward}

\begin{algorithmic}
\REQUIRE Network with $L$ layers - $0,1,2,....,L-2,L-1$
\REQUIRE Set of weighting parameters - $W^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of bias parameters - $b^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of layer activation function - $a^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Input sample - $x^{(0)}$

$Z = \{\} \leftarrow$ array to hold pre-activations \\
$X = \{\} \leftarrow$ array to hold final activations

\FOR {$l = 1,2,...,L-2,L-1$}
	\item Compute pre-nonlinear activation
	\item $z^{(l)} \leftarrow W^{(l)}x^{(l-1)} + b^{(l)}$
	\item Apply activation function
	\item $x^{(l)} \leftarrow a^{(l)}[z^{(l)}]$
	\item Store values for later use in optimizer
	\item $Z \leftarrow Z.add(z^{(l)})$
	\item $X \leftarrow X.add(x^{(l)})$
\ENDFOR

\item Network final prediction is last object in 'act' array.
\item $y^* \leftarrow x^{(L-1)} = X[-1]$
\item Return the pre-nonlinear activations and final-nonlinear activations
\RETURN $Z , X$

@@

\subsection{Layers Used in Classification Neural Network}
\label{sec-Layers}

@@

\paragraph*{}Typically, a layer function can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries. Step (i.) is usually in the form of a matrix-vector equation:
\begin{equation}
\label{eqn-matVectEqn}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
Where $W^{(l)}$ is the \textit{weighting-matrix} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.
Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-matVectEqn}). $\sigma^{(l)}$ is some activation function, which is often use to enable the modeling of more complex-decision boundaries.

\paragraph*{}The combination of eqn. (\ref{eqn-matVectEqn}) and eqn. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in eqn. (\ref{eqn-altLayerFunction}). Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

\subsubsection{Dense Layer}
\label{subsec-DenseLayer}

@@

\paragraph*{}We model a layer of neurons as a vector of floating-point numbers. Typically, it is required that the array be one-dimensional. Suppose a layer $(l)$ contains $n$ artificial neurons. We denote the array that hold those activations as $x^{(l)}$ and is given by:
\begin{equation}
\label{layer-DenseNeurons}
\vec{x}^{(l)} = \Big[ x_0, x_1, x_2, \hdots , x_{n-2}, x_{n-1} \Big]^T
\end{equation}
The activation of each entry is given by a linear-combination of activations from the previous layer, as outlined in eqn.(\ref{eqn-matVectEqn}) and eqn.(\ref{eqn-elementActivation}).

@@

\subsubsection{2-Dimensional Convolution Layer}

@@

@@

\paragraph*{}The 1-Dimensional Concatenation Layer, also called 1D-Concat layer takes separate vectors of activations

@@

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{sec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in eqn. (\ref{eqn-matVectEqn}), and then an element-wise activation function as in eqn. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in out classification neural network.

@@

\paragraph*{}The Rectified Linear Unit (ReLU) activation function acts element-wise on the activations in a given layer. If the activation of a neurons is non-zero or non-negative, the value is untouched, otherwise a $0$ is returned. For an input activation $x$, ReLU is defined by:
\begin{equation}
\label{eqn-ReLU}
\text{ReLU}\big[x\big] = \max\big(0,x\big) = \left\{
\begin{array}{ll}
    x & \text{if\hspace{0.5cm}} x > 0 \\
    0 & \text{otherwise}
\end{array} 
\right.
\end{equation}
ReLU is only defined on a real domain. We provide a visualization of the function in fig. (\ref{fig-ReLU}).
\begin{figure}[H]
\label{fig-ReLU}
\begin{center}
\includegraphics[scale=0.2]{../Figures/RectifiedLinearUnit}
\end{center}
\caption{Rectifed Linear Unit (ReLU) activation function}
\end{figure}

@@

@@

\subsection{Training a Neural Network}
\label{sec-Training}

@@

\paragraph*{}The success of a neural network algorithm is enormously dependent of the strength of the chosen features, rather than the quantity of available data \cite{Geron,Kahn,Liu}. We have derived an appropriate set of inputs arrays \textcolor{blue}{See Section on Features}, and constructed a neural network architecture to compliment these features. From these two input arrays, one being a matrix and the other being a column-vector, we have produced a \textit{multimodal neural network}, that process the two inputs, and combines separate neural networks into a single model to produce a one prediction resulting from both inputs.

@@

\paragraph*{}By grouping layers in this structure, each of the $32$ filters passes over a $3 \times 3$ area, and then again. The result is

@@

\paragraph*{} The Perceptron branch is picture on the right side of fig. (\ref{fig-NetworkArchitecture}), notice that it is considerably smaller in side and complexity than it's neighbor. Rather than accept an image-like input, the perception simply takes a vector-like input of properties of a larger data set - in this case, each waveform. We call these properties \textit{features} \cite{Geron,Kahn,Virtanen}. \textcolor{blue}{Details on the creating of these features can be found in the \textit{features} section of this document.}

\paragraph*{}The input layer of the perceptron accepts a 2-Dimensional array. The axes, in order of indexing, represent (i) the size of the \textit{mini-batch} of samples, (ii) the number of features for each sample. We use the same model hyper-paramater of $64$ samples per batch, and have developed $23$ unique clssification features that have been derived from time-space and frequency-space representations of the audio file data. We can denote the 2D shape if the input object into this branch as:
\begin{equation}
\label{eqn-shapeX2}
X_2 \in \mathbb{R}^{(64 \times 20)}
\end{equation}
Again, any other shape will be rejected by the model, and an error is raised.

@@

\paragraph*{}The last layer in eachof the two branches is a ReLu-activated Dense layer containing 64 neurons. The activations are concatenated

@@

>>>>

% ================================
% Landon Buell
% Kevin Short
% Physics 799%
% 26 September 2020
% ================================

@@

% ================================

\DeclareMathOperator{\sign}{\text{sign}}

% ================================

@@

% ================================================================
% ================================================================

\begin{titlepage}
	\begin{center}
   		\vspace*{1cm}
       	\begin{LARGE}
       	\textbf{Classifying Chaotic Synthesizers with a Multimodal Neural Network} \\
       	\end{LARGE}
       	\vspace{1cm}
       	\begin{Large}
       	Thesis Subtitle 
       	\end{Large}             
       	\vspace{1.5cm}

       	\textbf{By: \\ Landon Buell$^1$} \\       
       	\vspace{0.5cm}
       	\textbf{Advised By: \\ Dr. Kevin Short$^2$ , Dr. Maurik Holtrop$^1$} \\
      	\vspace{1cm}
      	A thesis presented for the degree of\\
       	Bachelor of Science in Physics
        \vspace{1cm}

       \vfill

       \vspace{1cm}              
       $^1$Department of Physics and Astronomy\\           
       $^2$Department of Mathematics and Statistics\\
       \vspace{1cm}  
       University of New Hampshire\\
       Durham New Hampshire, USA\\
       Dec 2020\\

   \end{center}
\end{titlepage}

\tableofcontents

% ================================================================
% ================================================================
\newpage
\section{Introduction}
\label{sec-Introduction}

% ================================================================
% ================================================================
\newpage
\section{Methodology}
\label{sec-Methodology}

\paragraph*{}We outline and detail the steps taken to execute this project from beginning to end.
\textcolor{red}{Does this section need to be included? I'm not sure how to format or organize it}

@@

\subsection{Designing the Function}

\paragraph*{}Consider the biological process of hearing a sound wave and matching it to a source. We can model this behavior by some unknown function $F$. We produce an approximation of that function $F^*$ that can map the contents of a sound file that have been generated by a chaotic synthesizer to a potential source. For a set of inputs $\vec{x} = \big\{ x_0 , x_1 , x_2 , ... , x_{p-1} \big\}$ and a set of classes $0$ through $k-1$, we denote this function as:
\begin{equation}
\label{eqn-MappingFunction}
F^*: \vec{x} \rightarrow \big\{ 0 , 1 , 2 , ... , k-1 \big\}
\end{equation}

@@

\subsection{Collecting and Pre-processing Raw Data}

\paragraph*{}In order to train the Multimodal neural network to identify musical instrument sources, we need a suitable data set to present to the model. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \textcolor{blue}{Citations!}. These contain short segments of musical instruments performing a single note or a collection of notes in succession.

\paragraph*{}To ensure that these data sets are roughly homogeneous, we read each sample from it's original format, \textit{.aif}, \textit{.mp3}, or similar, and rewrite each sample as a new \textit{.wav} files, sampled at 44.1 kHz, with a 16 bit depth \textcolor{red}{check this!}. This ensures that all data will have a consistent format when features are extracted. We also use this stage to ensure that each audio file has a correct label, and to determine the number of unique output classes.

\subsection{Designing Classification Features}

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of classification features. These are properties of wave forms that can be represented by a numerical value or several numerical values and are used as the primary tool in classification. There are used in place of a full waveform to represent a file's contents We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

\subsection{Designing A Complementary Network Architecture}

\paragraph*{}With the appropriate set of features designed, and the number of output determined, we can organize the structure of the neural network function. We construct a multimodal neural network than processes two input arrays derived from the same audio sample that share a common label. This network is designed to handle and process each respective input independently, and concatenate the results to produce a single output.

\subsection{Testing and Evaluating Network Performance}

\paragraph*{}We divide the raw data set up into a training and testing sets. We employ cross-validation and compute the results of performance metrics to ensure that out model is making reliable predictions, and generalize appropriately. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance.

\subsection{Running Predictions of Chaotic Synthesizer Files}

\paragraph*{}Once we have established a suitable performance of the model, we allow the model to run predictions on the un-labeled chaotic synthesizer wave forms. We output the prediction results to a file, and compare the neural network predictions against human predictions. If further corrections are needed, we revert and re-deign the features, architecture, or hyper-parameters as needed.

% ================================================================
% ================================================================

\newpage
\section{The Neural Network}
\label{sec-TheNeuralNetwork}

\subsection{An Introduction to Neural Networks}
\label{subsec-NerualNetworkIntro}

@@

\subsection{The Structure of a Neural Network}
\label{subsec-NetworkStructre}

\paragraph*{}A Neural Network is simply a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. This output encodes the models final "decision" given a unique input.

@@

\paragraph*{}The model is recursive by nature, the activations from one layer, $l-1$, are used to directly produce the activations of the next successive layer $l$. Thus eqn. (\ref{eqn-LayerFunction}) can be alternatively written as:
\begin{equation}
\label{eqn-altLayerFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R} \rightarrow x^{(l)} \in \mathbb{R}
\end{equation}
The array of activations, $x^{(0)}$, is the raw input given the neural network, called \textit{input} activations. Conversely, $x^{(L-1)}$ is referred to as\textit{output} activations \cite{Geron,James,Loy}.

\begin{algorithm}[H]
\caption{Forward propagation system in a standard deep neural network. Each iteration in the main \textit{for-loop} represents the execution of a layer, and passing the result to the "next" layer function. We assume each layter to follow a two-step structure, being (i) a linear transformation as in eqn. (\ref{eqn-LinearTransform}) and (ii) an element-wise non-linear activation function as in eqn. (\ref{eqn-elementActivation}).  A practical application of this algorithm should include batches of samples instead of a single sample.
\textcolor{red}{I Want to include this algorithm for context - especially on the programming end, but am not sure where to put it. }}
\label{algFeedForward}

\begin{algorithmic}
\REQUIRE Network with $L$ layers - $0,1,2,....,L-2,L-1$
\REQUIRE Set of layer functions $f^{(l)}, l \in 0,1,2,....,L-2,L-1$ made up of 
\REQUIRE Input sample - $x^{(0)}$

$Z = \{\} \leftarrow$ array to hold pre-activations \\
$X = \{\} \leftarrow$ array to hold final activations \\

\FOR {$l = 1,2,...,L-2,L-1$}

	\item Break layer function into two linear transformation and non-linear activation

	\item Store values for later use in optimizer
	\item $Z \leftarrow Z.add(z^{(l)})$
	\item $X \leftarrow X.add(x^{(l)})$
\ENDFOR

\item Network final prediction is last object in 'act' array.
\item $y^* \leftarrow x^{(L-1)} = X[-1]$
\item Return the pre-nonlinear activations and final-nonlinear activations.\\
\item The final index of $X$ represents the neural network's final prediction. \\
\RETURN $Z , X$

@@

\subsection{Layers Used in Classification Neural Network}
\label{subsec-Layers}

@@

\paragraph*{}Typically, a layer function can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries. Step (i.) is usually in the form of a matrix-vector equation:
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
Where $W^{(l)}$ is the \textit{weighting-matrix} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.
Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-LinearTransform}). $\sigma^{(l)}$ is some activation function, which is often use to enable the modeling of more complex-decision boundaries.

\paragraph*{}The combination of eqn. (\ref{eqn-LinearTransform}) and eqn. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in eqn. (\ref{eqn-altLayerFunction}). Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

% ================================

\subsubsection{Dense Layer}
\label{subsubsec-DenseLayer}

@@

\paragraph*{}We model a layer of neurons as a vector of floating-point numbers. Typically, it is required that the array be one-dimensional. Suppose a layer $(l)$ contains $n$ artificial neurons. We denote the array that hold those activations as $x^{(l)}$ and is given by:
\begin{equation}
\label{layer-DenseNeurons}
\vec{x}^{(l)} = \Big[ x_0, x_1, x_2, \hdots , x_{n-2}, x_{n-1} \Big]^T
\end{equation}
The activation of each entry is given by a linear-combination of activations from the previous layer, as outlined in eqn.(\ref{eqn-LinearTransform}) and eqn.(\ref{eqn-elementActivation}).

@@

% ================================

\subsubsection{2-Dimensional Convolution Layer}
\label{subsubsec-Conv2DLayer}

@@

% ================================

@@

\paragraph*{}The 1-Dimensional Concatenation Layer, also called 1D-Concat layer takes separate vectors of activations and combines them into a single 1D vector. Consider the layer activations

@@

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{sec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in eqn. (\ref{eqn-LinearTransform}), and then an element-wise activation function as in eqn. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in out classification neural network.

@@

\paragraph*{}The Rectified Linear Unit (ReLU) activation function acts element-wise on the activations in a given layer. If the activation of a neurons is non-zero or non-negative, the value is untouched, otherwise a $0$ is returned. For an input activation $x$, ReLU is defined by:
\begin{equation}
\label{eqn-ReLU}
\text{ReLU}\big[x\big] = \max\big(0,x\big) = \left\{
\begin{array}{ll}
    x & \text{if\hspace{0.5cm}} x > 0 \\
    0 & \text{otherwise}
\end{array} 
\right.
\end{equation}
ReLU is only defined on a real domain. We provide a visualization of the function in fig. (\ref{fig-ReLU}).
\begin{figure}[H]
\label{fig-ReLU}
\begin{center}
\includegraphics[scale=0.2]{../Figures/RectifiedLinearUnit}
\end{center}
\caption{Rectifed Linear Unit (ReLU) activation function}
\end{figure}
For our chosen architecture, ReLU is applied to the activations in every single Convolution layer and Dense Layer, with the exception of the output dense layer.

% ================================

@@

\paragraph*{}The softmax activation function is commonly used in the output layer of a multi-class classification network, such as the one we implement. When softmax acts on a vector or activation, the result is a normalized,non-negative output vector. \cite{Geron,Goodfellow,Virtanen}. The $i$-th element in a softmax activation function is given by:
\begin{equation}
\label{eqn-Softmax}
\text{Softmax}\big[ x \big]_{i} = \frac{1}{\sum_{j}e^{x_j}}e^{x_i}
\end{equation}

\paragraph*{}This function is only used in the output layer of a neural network, and encodes a discrete categorical probability. For example, if eqn.(\ref{eqn-Softmax}) returns a vector of the form: $[ 0.75 , 0.25]$, we interpret this as a sample having a $75\%$ chance of beloinging to class $0$ and a $25\%$ change of belonging to class $1$.

@@

\subsection{Training a Neural Network}
\label{subsec-Training}

@@

\paragraph*{}The success of a neural network algorithm is enormously dependent of the strength of the chosen features, as well as a complementary architecture, rather than the quantity of available data \cite{Geron,Kahn,Liu}. We have derived an appropriate set of inputs arrays \textcolor{blue}{See Section on Features}, and constructed a neural network architecture to coincide with these features. From these two input arrays, one being a matrix and the other being a column-vector, we have produced a \textit{multimodal neural network}, that process the two inputs, and combines separate neural networks into a single model to produce a one prediction resulting from both inputs \textcolor{blue}{citation for multimodal networks}.

@@

\paragraph*{}By grouping layers in this structure, we use the convolution  layers to reduce the number of features, and then the pooling layer to extract only the largest activations of the remianing features. This ensures that in the full spectrogram image, only the dominant features and feature shapes are preserved and passed into the next layer set for processing.

@@

\paragraph*{} The Perceptron branch is picture on the right side of fig. (\ref{fig-NetworkArchitecture}), notice that it is considerably smaller in side and complexity than it's neighbor. Rather than accept an image-like input, the perception simply takes a vector-like input of properties that are derived from the audio file. We call these properties \textit{features} \cite{Geron,Kahn,Virtanen}. \textcolor{blue}{Details on the creating of these features can be found in the \textit{features} section of this document.}

\paragraph*{}The input layer of the perceptron accepts a 2-Dimensional array. The axes, in order of indexing, represent (i) the size of the \textit{mini-batch} of samples, (ii) the number of features for each sample. We use the same model hyper-paramater of $64$ samples per batch, and have developed $20$ unique classification features that have been derived from time-space and frequency-space representations of the audio file data. We can denote the 2D shape if the input object into this branch as:
\begin{equation}
\label{eqn-shapeX2}
X_2 \in \mathbb{R}^{(64 \times 20)}
\end{equation}
This 2D is referred to as a \textit{design matrix} \cite{James,Loy}. Any other shape will be rejected by the model, and an error is raised.

\paragraph*{}In perceptron models, scaling the design matrix appropriately drastically improves the classification performance. A design matrix is scaled by taking all samples in a column (a particular feature from each class), subtracting the average, and then scaling it such that it has unit variance \cite{Geron,James}. This ensures that no one feature dominates the performance of the model and that layer activations to not get saturated as data progresses through the network.

@@

\paragraph*{}The last layer in each of the two branches is a ReLU-activated Dense layer containing 64 neurons, represented by a $64 \times 1$ column vectors. We combine these model layers by using a \textit{concatenation layer} to fuse the two column vectors together, "vertically", such that the result is a $128 \times 1$ column vector. This vector is then passed into the final dense layer, which used the softmax activation, and encodes the joint predictions based on both model branches.

@@

\section{Feature Selections}
\label{sec-FeatureSelections}

\paragraph*{}Even best minds in the world will perform poorly on test which they have studied the wrong material- neural networks are no different. In order to properly train a neural network, the model must be presented with an appropriate set of training input $x$ and a complementary set of training labels $y$ \cite{Geron,Goodfellow,James}. It becomes quickly apparent that the nature of information contained in the input object $x$, called \textit{features}, is \textit{extremely important} to the performance of the classifier. Consider if you were tasked to identify cats and dogs from images, but instead were presented with only the top-most row of pixels- The task would be nearly impossible because of an inappropriate or in complete set of information.

\paragraph*{}Tuomas Virtanen, machine learning and audio engineer writes in his book, "Computational Analysis of Sound Scene and Events" \cite{Virtanen}:
\begin{quote}
For recognition algorithms, the necessary property of the acoustic features is low variability among features extracted from examples assigned to the same class, and at the same time high variability allowing distinction between features extracted from examples assigned to different classes.
\end{quote}
In constructing a neural network classifier, the development of appropriate features is of the utmost importance. To ensure the construction of a suitable model, we derive features based from three sources (i) a spectrogram matrix of the waveform, (ii) the time-space representation of the waveform, and (iii) the frequency space representation of the waveform. It is important to note that although this algorithm will classify sound waves to instruments, the model will never actually be presented with a waveform directly, instead it will rely on these features.

\paragraph*{}Once we produce an sufficient set of features, we concatenate them into a single object, $\hat{x}$, called the \textit{feature-vector} \cite{Goodfellow}. In the training process, this object, along with the appropriate classification label, $y$ is presented to the neural network for processing. This process of constructing a feature vector from any data set is vital and is used to represent the data set in a far more compact and non-redundant format \cite{Virtanen,Liu}

\paragraph*{}To ensure suitable performance of this sound wave classification neural network, a great deal of time has been devoted to the construction of the elements of the feature vector. These features are derived are principles of music, digital signal processing, previous work success, and most importantly physics. In the following sections, we outline the set of $21$ features used in the classification process

% ================================

\subsubsection{The Design Matrix}

\paragraph*{}The $21$ features are from each file are organized into one of two array structures. each called a \textit{design matrix}. A design matrix is an N-Dimensional array of real floating-point numbers that stores the features in a convenient way for the neural network to digest. For this hybrid neural-network, we use two matrices, $X_1$ and $X_2$, each containing data from $b$ audio file samples. each has shape given by:
\begin{equation}
\label{eqn-X1Shape}
X_1 \in \mathbb{R}^{(b \times N' \times k' \times 1)}
\end{equation}
\begin{equation}
\label{eqn-X2Shape}
X_2 \in \mathbb{R}^{(b \times p)}
\end{equation}
Where $N' \times k' \times 1$ gives the shape of a single spectrogram matrix $S$, described in sect. (\ref{subsec-spectrogram}), The first axis, with length $b$ indicates that there are $b$ samples stored in that design matrix.
Similarly, $p$ is the number of features derived from the time and frequency-representation of features outlined in sec. (\ref{subsec-time}) and sec.(\ref{subsec-frequency}). A row of $X_2$ has shape $1 \times p$ where there are $p$ features within the \textit{feature-vector} for each sample. Matrix $X_1$ is presented to the convolutional branch of the network to be process as an a collection of images, and matrix $X_2$ is presented to the perceptron branch to be processed as a collection of vectors.

% ================================

\subsubsection{Audio Preprocessing}

\paragraph*{}Preprocessing a data set is a necessary step to execute prior to feature extraction \cite{Geron2,James}. In the case of audio files, preprocessing usually consists of ensuring that the data set contains the following:
\begin{enumerate}
\item A suitably sized number of files, of reasonable audio quality, with normalized amplitudes
\item Audio encoded in a standard, and consistent format
\item A consistent sample rate and bit depth between audio files
\item A consistent number of channels
\end{enumerate}
Note that different projects may require a different set of requirement from preprocessing \cite{Virtanen}. For this project, we have chosen to use the following parameters:
\begin{enumerate}
\item Roughly $4000$ audio files Professionally or semi-professionally recorded in a studio. \textcolor{red}{Citation needed!}. All ampltiudes have been normalized to $\pm 1$ unit.
\item All audio has be converted into \textit{.WAV} files from other formats, such as \textit{.AIF} or \textit{.MP3} using a MATLAB program
\item All audio is sampled at $44,100$ Hz and \textcolor{blue}{Bit depth needed - 16? 24?}
\item All audio has been down-mixed into mono-channel waveforms.
\end{enumerate}

\subsection{Spectrogram Feature}
\label{subsec-spectrogram}

\paragraph*{}The field of neural classification is well studied in the application of image-processing. Many large-scale, and even introductory neural network projects find themselves under the umbrella of image classification \cite{Geron,Goodfellow,Loy,Mierswa}. As a result, model architectures for image processing related tasks are well-explored and have shown experimentally successful behavior. Following this success, it make senses to provide an image-like representation of a sound wave as a feature. We do this in the form of a spectrogram matrix.

\paragraph*{}A spectrogram is a representation of the energy distribution of a sound wave as a function of both space and time. In a conventional spectrogram, the passing of time is shown along the $x$-axis, and the frequency spectrum is shown on the $y$-axis. Thus each point in the 2-Dimensional space is an energy at a given time and frequency. Examples spectrograms from the wave form data set are shown in Fig. (\ref{fig-spectrograms}).

\begin{figure}[H]
\begin{center}
\label{fig-spectrograms}
\textcolor{red}{Insert spectrograms here}
\caption{Spectrogram representations of various waveforms}
\end{center}
\end{figure}

\paragraph*{}A spectrogram is produced by the method of \textit{frame-blocking}, which is very prevalent in audio signal classification. Frame-blocking takes a raw waveform or signal, $s$ and decomposes it into a set of analysis frames, $a_i$, with each being $N$ samples in length, and has a fixed overlap with the next adjacent frame. Each of the $k$ frames then allows for a section of the signal in somewhat stationary state \cite{Liu,Zhang,Kahn,Virtanen}. For this project, we have chosen to use frames of size $N = 4096$ with a $75\%$ or $3072$ sample overlap. At a sample rate of $f_s = 44100$ samples/second, each frame represents a slice of time that is about $0.1$ seconds long.

\paragraph*{}We concatenate each analysis frame, $a_{i}, i \in [0,k-1]$ into a single $k \times N$ matrix, called $A$. Each row is a frame, each column is an index in that frame
\begin{equation}
\label{eqn-FrameMatrix}
A = \big\{ a_{0} , a_{1} , a_{2} , ... , a_{k-1} \big\} = 
\begin{bmatrix}
a_{0}[0] & a_{0}[1] & a_{0}[2] & \hdots & a_{0}[N-1] \\
a_{1}[0] & a_{1}[1] & a_{1}[2] & \hdots & a_{1}[N-1] \\
a_{2}[0] & a_{2}[1] & a_{2}[2] & \hdots & a_{2}[N-1] \\
\vdots 		& \vdots 	  & \vdots 		& \ddots & \vdots        \\
a_{k-1}[0] & a_{k-1}[1] & a_{k-1}[2] & \hdots & a_{k-1}[N-1] \\
\end{bmatrix}
\end{equation}
We use bracket notation, $[j]$ to indicate that each analysis frame $a_i$ is array-like. Many programming languages could also use the following indexing conventions for matrix $A$:
\begin{equation}
\label{eqn-IndexingA}
A_{i,j} = A_i[j] = A[i][j] = A[i,j]
\end{equation}
These all represent equivalent entries.

\paragraph*{}After frame-blocking, we apply a \textit{windowing function} to each frame. A standard \textit{Hann Window} of $N$ samples is generated and reshaped into a $N \times 1$ column-array, $H$. The $n$-th index in a Hann window with $N$ samples is defined:
\begin{equation}
\label{eqn-Hann}
H[n] = \frac{1}{2}\bigg[ 1 - \cos\Big( \frac{2\pi n}{N-1}\Big)\bigg]
\end{equation}
The window function is applied to each analysis frame by computing the element-wise product of the Hann window array, $H$ and each row of the analysis frames matrix, $A_{i,:}$. Result is an $k \times N$ array, $\widetilde{A}$:
\begin{equation}
\label{eqn-WindowMatrix}
\widetilde{A_i} = A_i \odot H
\end{equation}

\paragraph*{}Finally, we perform a \textit{Discrete Fourier Transform} (DFT) to bring each analysis frame from a time domain into a frequency domain \cite{Olson,Peatross}. The Discrete Fourier Transform is applied by producing an $N \times N$ \textit{transform matrix}, often noted at $\mathbb{W}$. Let $\omega^k = e^{\frac{-2\pi i}{N}k}$, then the DFT matrix for a time-space containing $N$ samples
\begin{equation}
\label{eqn-DFTMatrix}
\mathbb{W} = \frac{1}{\sqrt{N}}
\begin{bmatrix}
1 & 1 & 1 & 1 & \hdots & 1 \\
1 & \omega		& \omega^2 & \omega^3 & \hdots & \omega^{N-1} \\
1 & \omega^2	& \omega^4 & \omega^6 & \hdots & \omega^{2(N-1)} \\
1 & \omega^3	& \omega^6 & \omega^9 & \hdots & \omega^{3(N-1)} \\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots \\
1 & \omega^{N-1} & \omega^{2(N-1)} & \omega^{3(N-1)} & \hdots & \omega^{(N-1)^2}
\end{bmatrix}
\end{equation}
Each column of the matrix is a complex sinusoidal oscillating with an integer number of periods within the $N$-sample length window \cite{Short,Peatross}.The DFT is applied by a taking the matrix -product of $\mathbb{W}$ and $\widetilde{A}^T$. The transpose of $\widetilde{A}$ then makes each analysis frame into a column vector, which gives the appropriate dimension for multiplication.

\paragraph*{}Most standard implementations of neural network models require all activations, weights, and biases to be real floating-point numbers. Since the the DFT matrix introduces complex values, we compute the square of the element-wise $L2$-norm of the resultant array.
\begin{equation}
\label{eqn-Spectrogram}
S_{i,j} = \big{\lVert} (\mathbb{W}\widetilde{A}^T)_{i,j} \big{\rVert}_{2}^{2}
\end{equation}
Where $\mathbb{W}$ is the DFT matrix from eqn. (\ref{eqn-DFTMatrix}) and $\widetilde{A}^T$ is the transpose of the analysis frames matrix from eqn. (\ref{eqn-WindowMatrix}). The matrix $S$, is the spectrogram representation of the initial waveform, and has shape $N \times k$ of real floating-point numbers. We can index matrix $S$ similarly to that of matrix $A$ in eqn. (\ref{eqn-IndexingA}):
\begin{equation}
\label{eqn-IndexingS}
S_{i,j} = S[i][j] = S[i,j]
\end{equation}

\paragraph*{}Each column of the $S$ matrix is now a single-frame that has been moved into a frequency-space representation, thus there are $k$ columns, just as there were $k$ time-series analysis frames. Given the discretized nature of digital audio, the frequency-space representation is not a continuous function, but rather a column vector, where the frequency has been assigned to one of $N$ bins. Since human hearing extends from $20$ Hz to $20$ kHz, audio files such as music and voice recordings are typically sampled around $44.1$ kHz, to ensure that the full range of audio is held by the recording \cite{Olson,Virtanen}.

\paragraph*{}However, standard musical instruments seldom extend above $8$ kHz \cite{Olson,Virtanen,White}. This means that when constructing the spectrogram, we will rarely ever see energy present above this frequency at any time, and the $S$ matrix will contain mostly zero, or zero-like entries. As a result of this, we can simply crop each matrix to only represent frequencies that are only between $0$ Hz and $8000$ Hz. This makes the input array smaller, and eliminates redundant and non-useful information. The number rows in the $S$ matrix is reduced from $N$ down to $N'$. Since each raw audio file may contain a different number of samples to begin with, a different number of analysis frames are created. To ensure all samples are a homogeneous size, we assert that each matrix must have exactly $k'$ columns. If $k' < k$, columns of zeros are added to pad the input, and if $k' > k$, columns are removed.
